\documentclass[11pt]{article}
\usepackage{fullpage}
\usepackage{pythonhighlight}

\usepackage[left=1in,top=1in,right=1in,bottom=1in,headheight=3ex,headsep=3ex]{geometry}
\usepackage[T1]{fontenc}
\newcommand{\blankline}{\quad\pagebreak[2]}
\usepackage{multirow}
\title{ECON 2250: Statistics for Economists}
\author{Jim Crozier}
\date{Fall 2022}

\usepackage[sc]{mathpazo}
\linespread{1.05}         % Palatino needs more leading (space between lines)
\usepackage[T1]{fontenc}
\newcommand{\emdash}{\nobreak-\nobreak\hskip0pt}
\usepackage[mmddyyyy]{datetime}% http://ctan.org/pkg/datetime
\usepackage{advdate}% http://ctan.org/pkg/advdate
\newdateformat{syldate}{\twodigit{\THEMONTH}/\twodigit{\THEDAY}}
\newsavebox{\MONDAY}\savebox{\MONDAY}{Mon}% Mon

\newcommand{\week}[1]{%
%  \cleardate{mydate}% Clear date
% \newdate{mydate}{\the\day}{\the\month}{\the\year}% Store date
  \paragraph*{\kern-2ex\quad #1, \syldate{\today} - \AdvanceDate[4]\syldate{\today}:}% Set heading  \quad #1
%  \setbox1=\hbox{\shortdayofweekname{\getdateday{mydate}}{\getdatemonth{mydate}}{\getdateyear{mydate}}}%
  \ifdim\wd1=\wd\MONDAY
    \AdvanceDate[7]
  \else
    \AdvanceDate[7]
  \fi%
}



\usepackage{setspace}
\usepackage{multicol}
%\usepackage{indentfirst}
\usepackage{fancyhdr,lastpage}
\usepackage{url}
\pagestyle{fancy}
\usepackage{hyperref}
\usepackage{lastpage}
\usepackage{amsmath}
\usepackage{layout}   
\lhead{}
\chead{}
\rhead{\footnotesize Statistics for Economics  -- Fall 2022}
\lfoot{}
\cfoot{\small \thepage/\pageref*{LastPage}}
\rfoot{}

\usepackage{array, xcolor}
\usepackage{color,hyperref}
\definecolor{clemsonorange}{HTML}{B3A369}
\hypersetup{colorlinks,breaklinks,
            linkcolor=clemsonorange,urlcolor=clemsonorange,
            anchorcolor=clemsonorange,citecolor=black}




\begin{document}


\maketitle

\blankline

Mid term practice 1

\vspace{10 mm}

Suppose X is a column in a panda's dataframe and $x_1, x_2, ..., x_n$ are the rows of X where n is the number of observations. For reference, the python code for the math we are about to type is.

\begin{python}
import pandas as pd
import numpy as np
df = pd.DataFrame({'X':[5, 7, 10]})
n_obs = len(df)
mean_x = df["X"].mean()
np.sqrt(sum((df["X"] - mean_x)**2)/(n_obs-1))
\end{python}

\subsection{jargony jargon}
Define the following words:

\begin{itemize}
  \item scalar
  \item vector 
  \item dataframe 
  \item sum
  \item quantile
  \item median
  \item 
\end{itemize}


\subsection{ Mean to Standard Deviation }

Use the summation operator ($\sum$) to express the sum of x from 1 to n. I just want the mathematical expression here.
\newline 
\newline
\newline 


 What does the scalar from above represent? 
\newline 
\newline
\newline 

Use the summation operator ($\sum$) to express the average of X. I just want the mathematical expression here.
\newline 
\newline
\newline 


 What does the scalar from above represent? 
\newline 
\newline
\newline 


Refer to the above scalar as \emph{mean$\_$x} and write an equation for subtracting off \emph{mean$\_$x} from element of X. Refer to the elements as $x_i$, and I only want the equation. 
\newline 
\newline
\newline 


What does the list (vector) from above represent? How many rows will it have?
\newline 
\newline
\newline 

Write an equation for squaring the above equation.
\newline 
\newline
\newline 


What does the list (vector) from above represent? How many rows will it have?
\newline 
\newline
\newline 


Write an equation for summing the squared elements above and dividing by n. This should look kind of similar to taking the average above.
\newline 
\newline
\newline 


Now take the square of the above equation. What do we call this observation, and what does it represent.
\newline 
\newline
\newline 



\subsection{  Quantiles : the sorting hat}

Describe how you would find the min and max, and what they represent. 
\newline 
\newline
\newline 

Describe how you find the median? What the core idea behind median and quantiles? What quanitle does the median represent? 
\newline 
\newline
\newline 

What is the difference in finding the median of an even or odd number of observations.
\newline 
\newline
\newline 


Why do we care about the difference between the mean or median or other quantiles? 
\newline 
\newline
\newline 


\end{document}
