\documentclass[11pt]{article}
\usepackage{fullpage}
\usepackage{pythonhighlight}

\usepackage[left=1in,top=1in,right=1in,bottom=1in,headheight=3ex,headsep=3ex]{geometry}
\usepackage[T1]{fontenc}
\newcommand{\blankline}{\quad\pagebreak[2]}
\usepackage{multirow}
\title{ECON 2250: Statistics for Economists}
\author{Jim Crozier}
\date{Fall 2022}

\usepackage[sc]{mathpazo}
\linespread{1.05}         % Palatino needs more leading (space between lines)
\usepackage[T1]{fontenc}
\newcommand{\emdash}{\nobreak-\nobreak\hskip0pt}
\usepackage[mmddyyyy]{datetime}% http://ctan.org/pkg/datetime
\usepackage{advdate}% http://ctan.org/pkg/advdate
\newdateformat{syldate}{\twodigit{\THEMONTH}/\twodigit{\THEDAY}}
\newsavebox{\MONDAY}\savebox{\MONDAY}{Mon}% Mon

\newcommand{\week}[1]{%
%  \cleardate{mydate}% Clear date
% \newdate{mydate}{\the\day}{\the\month}{\the\year}% Store date
  \paragraph*{\kern-2ex\quad #1, \syldate{\today} - \AdvanceDate[4]\syldate{\today}:}% Set heading  \quad #1
%  \setbox1=\hbox{\shortdayofweekname{\getdateday{mydate}}{\getdatemonth{mydate}}{\getdateyear{mydate}}}%
  \ifdim\wd1=\wd\MONDAY
    \AdvanceDate[7]
  \else
    \AdvanceDate[7]
  \fi%
}



\usepackage{setspace}
\usepackage{multicol}
%\usepackage{indentfirst}
\usepackage{fancyhdr,lastpage}
\usepackage{url}
\pagestyle{fancy}
\usepackage{hyperref}
\usepackage{lastpage}
\usepackage{amsmath}
\usepackage{layout}   
\lhead{}
\chead{}
\rhead{\footnotesize Statistics for Economics  -- Fall 2022}
\lfoot{}
\cfoot{\small \thepage/\pageref*{LastPage}}
\rfoot{}

\usepackage{array, xcolor}
\usepackage{color,hyperref}
\definecolor{clemsonorange}{HTML}{B3A369}
\hypersetup{colorlinks,breaklinks,
            linkcolor=clemsonorange,urlcolor=clemsonorange,
            anchorcolor=clemsonorange,citecolor=black}




\begin{document}


\maketitle

\blankline

Mid term practice 1

\vspace{10 mm}

Suppose X is a column in a panda's dataframe and $x_1, x_2, ..., x_n$ are the rows of X where n is the number of observations. For reference, the python code for the math we are about to type is.

\begin{python}
import pandas as pd
import numpy as np
df = pd.DataFrame({'X':[5, 7, 10]})
n_obs = len(df)
mean_x = df["X"].mean()
np.sqrt(sum((df["X"] - mean_x)**2)/(n_obs-1))
\end{python}

\subsection{summary stats}
\begin{itemize}
\item what is the sum of the list above (5,7,10)?
\item how many observations in the list?
\item what is the mean of the list?
\item demean each row (subtract off the mean), show your work to one decimal point. 
\item square the demeaned values, show your work. 
\item get the sum of squared values, show your work. 
\item divide this by n (number of observations). 
\item just put a square root over it (no need to calculate the sqrt), what is this value called and why does it matter? 
\end{itemize}
\subsection{jargony jargon}
Define the following words:

\begin{itemize}
  \item scalar
  \item vector 
  \item dataframe 
  \item column
  \item row
  \item continuous variable, give example 
  \item discrete variable, give example 
  \item dependent variable, give example 
  \item what is pandas, how do you load it into python
  \item what the sum of a column (vector)
  \item mean
  \item quantile
  \item median
  \item summation operator
  \item marginal probability 
  \item bayes rule
  \item $P(A \cup B)$
  \item $P(A \cap B)$
  \item $P(A , B)$
  \item $P(A | B)$
  \item independent events, what does it imply wrt $P(A | B)$
  \item mutually exclusive event
  \item what is a type error?
\end{itemize}


\subsection{ Mean to Standard Deviation }

Use the summation operator ($\sum$) to express the sum of x from 1 to n. I just want the mathematical expression here.
\newline 
\newline
\newline 


 What does the scalar from above represent? 
\newline 
\newline
\newline 

Use the summation operator ($\sum$) to express the average of X. I just want the mathematical expression here.
\newline 
\newline
\newline 


 What does the scalar from above represent? 
\newline 
\newline
\newline 


Refer to the above scalar as \emph{mean$\_$x} and write an equation for subtracting off \emph{mean$\_$x} from element of X. Refer to the elements as $x_i$, and I only want the equation. 
\newline 
\newline
\newline 


What does the list (vector) from above represent? How many rows will it have?
\newline 
\newline
\newline 

Write an equation for squaring the above equation.
\newline 
\newline
\newline 


What does the list (vector) from above represent? How many rows will it have?
\newline 
\newline
\newline 


Write an equation for summing the squared elements above and dividing by n. This should look kind of similar to taking the average above.
\newline 
\newline
\newline 


Now take the square of the above equation. What do we call this observation, and what does it represent.
\newline 
\newline
\newline 



\subsection{  Quantiles : the sorting hat}

Describe how you would find the min and max, and what they represent. 
\newline 
\newline
\newline 

Describe how you find the median? What the core idea behind median and quantiles? What quanitle does the median represent? 
\newline 
\newline
\newline 

What is the difference in finding the median of an even or odd number of observations.
\newline 
\newline
\newline 


Why do we care about the difference between the mean or median or other quantiles? 
\newline 
\newline
\newline 


\subsection{Probability}


Explain the rule for finding $P(A \cup B)$. Should this be larger than P(A)? Could it be the same size as P(A), if so how? 
\newline  
\newline
\newline 



Explain the the term "mutually exclusive" in your own words with regards to $P(A \cup B)$. 
\newline 
\newline
\newline 

What is the $P(A \cup B)$ of two mutually exclusive events with P(A) = 0.1 and P(B) 0.2? Draw a venn diagram to show this. 
\newline 
\newline
\newline 

Referring to the above question, now what if I tell you that the $P(A|B)$ is 0.05, now what the $P(A \cup B)$?  Draw a venn diagram to show this. Draw a venn diagram to show this. 
\newline 
\newline
\newline 



Explain the rule for finding $P(A \cap B)$. Should this be larger than P(A)? Could it be the same size as P(A), if so how? 
\newline  
\newline
\newline 


What is the $P(A \cap B)$ of two independent events with P(A) = 0.1 and P(B) 0.2? Draw a venn diagram to show this. 
\newline 
\newline
\newline 



Ok, now for some bayes. You might find these familiar. 75$\%$ of the children in Exton school have a dog, and 30$\%$ have a cat. Given that 60$\%$ of those that have cat also have a dog, what percent of those that have a dog also have a cat?
\newline 
\newline
\newline 




In a factory, machine X produces 60$\%$ of the daily output and machine Y produces 40$\%$ of the daily output.

2$\%$of machine X's output is defective, and 1.5$\%$ of machine Y's output is defective.

One day, an item was inspected at random and found to be defective. What is the probability that it was produced by machine X?


\subsection{Correlation}


What is the upper and lower bounds of correlation? 
\newline  
\newline
\newline 

If we are take the row wise deviations of x and y, explain in words how the algorithm enforces these bounds? 
\newline  
\newline
\newline 


Explain in words what a correlation of 0.8 means? 
\newline  
\newline
\newline 


Explain in words what a correlation of -0.8 means? 
\newline  
\newline
\newline 




Explain in words what a correlation of -0.08 means? 
\newline  
\newline
\newline 



Explain in words what a correlation of 0.08 means? 
\newline  
\newline
\newline 


We often say that "correlation is not causation",  briefly explain what that means (breifly = a few sentences)? 
\newline  
\newline
\newline 

\newpage

Use the data below the following questions. 

\begin{center}
\begin{tabular}{ |c|c|} 
 \hline
 y & x  \\ 
 \hline
 5 & 3  \\ 
 3 & 8  \\ 
 7 & 1  \\ 
 1 & 4  \\ 
 \hline
\end{tabular}
\end{center}

What is the mean of y and x? 
\newline  
\newline
\newline 

Create a grid below and demean x and y, and then square those columns. 
\newline  
\newline
\newline 
\newline  
\newline
\newline 
\newline  
\newline
\newline 

Sum the squares of those two columns (deamned x sauared, and demeaned y squared), and show the square root (no need to calculate) of each multiplied by each other. 
\newline 
\newline  
\newline
\newline 


In your grid create new column and calculate $ (x_i - \bar{x})(y_i - \bar{y})$ where $\bar{x}$ is the mean and $\bar{y}$ is the mean of y  . Note that this will n long. 
\newline 
\newline  
\newline
\newline 

Now sum up the numerator. 
\newline 
\newline  
\newline
\newline 

If the you were able to calculate the numerator and denomenator and the correlation came out to -0.61394061, what would be the interpretation? 
\newline 
\newline  
\newline
\newline 


Draw an y axis and an x axis and plot the points. Does the correlation make sense. 
\newline 
\newline  
\newline
\newline 
\newline 
\newline  
\newline
\newline 




\end{document}
